random forest and lme

### 2020-04-09 src/rf.Rmd

---
title: "random forest"
output: github_document
---

```{r}
suppressPackageStartupMessages(library(randomForest))
suppressPackageStartupMessages(library(tidyverse))
```

@@

```{r}
rf.mill <- randomForest(
  mill.rate ~  tax.class + municipality + total.assessment + past.mill,na.action = na.omit,
  data=train
)
```

```{r}
predict(rf.mill, test)
```

>>>>

---
title: "random forest and lme"
output: github_document
---

```{r}
suppressPackageStartupMessages(library(randomForest))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(nlme))
suppressPackageStartupMessages(library(ROCR))
```

@@

# Random Forest
```{r}
set.seed(0)

rf.mill <- randomForest(
  mill.rate ~  tax.class + municipality + total.assessment + past.mill,na.action = na.omit, mtry = 4,
  data=train, ntree=500
)
#Evaluate variable importance
importance(rf.mill)
varImpPlot(rf.mill)

rf.as <- randomForest(
  next.assess ~  tax.class + municipality + total.assessment + mill.rate,na.action = na.omit,mtry = 4,
  data=train
)

importance(rf.as)
varImpPlot(rf.as)

yhat.bag <- predict(rf.mill,newdata=test)
plot(yhat.bag, test$mill.rate, xlab="Predicted Mill Rate Using Test Set", ylab="Actual Mill Rate")
abline(0,1)

yhat.bag1 <- predict(rf.as,newdata=test)
plot(yhat.bag1, test$next.assess, xlab="Predicted Asssessment Value Using Test Set", ylab="Actual Assessment Value")
abline(0,1)
```

# Linear Mixed Effect Model
```{r}
# both random slope and random intercept
# different rate of change of assessment value and mill rate as well as initial assessment value and mill rate for each municipality 
lme.mill <- lme( mill.rate ~  tax.class + municipality + total.assessment + past.mill,na.action = na.omit, data=train, random = ~ 1 + total.assessment + past.mill | municipality, method = "ML")

summary(lme.mill)

lme.as <- lme(next.assess ~  tax.class + municipality + total.assessment + mill.rate ,na.action = na.omit, data=train, random = ~ 1 + total.assessment + mill.rate | municipality, method = "ML")

summary(lme.as)
```
